\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-08-27 paper/paper.tex
Working on bib.
Modify caption for drive-model 4 figure
Fixing references to ASA style.

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \cite{nelson}.  A common assumption in series systems is the time to failure for each risk is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as ``masking''.\\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et al. performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure \cite{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

@@

\section{Motivating Example}
\subsection{Backblaze hard drive data}
\label{sec:Data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is permanently removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its data publicly available through their website \cite{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.  No other analysis or modeling of the failure data is provided.  Backblaze does, however, encourage others to further analyze its data.

@@

\subsection{Single Drive-Model Analysis}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, drive-model 14.  The Kaplan-Meier estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures till about 10,000 hours, increases from about 10,000 to 20,000 hours, and then starts to gradually flatten out.  In addition, computer hard drives are known to have a mixture of early and late failures \cite{chan}.  Therefore, there is both an empirical and theoretical justification to apply the GLFP model.

@@

Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a detailed description).

In Figure 3 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.  The solid stepwise curve corresponds to the observed failures.  Censored drives are not plotted.  Standard error bands are again calculated using Greenwood's method \cite{green}.  Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.  The non-parametric and parametric model align up quite closely.  Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.  Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution. The discrepancy in the right tail of the distribution appears to be real. We don't provide a solution to this problem here. Certainly, if there are clearly more that two well-separated failure modes, the GLFP will not fit the data well. However, we consider it satisfactory if we can provide a good approximation to the observed lifetime distribution over the bulk of the product life. In practical terms, the right tail is of less importance for electronic goods, due to obsolescence.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\cite{betancourt}. HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density. This is done to reduce autocorrelation and efficiently explore the posterior. Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1,2 and 3 and 16 chains were run for Model 4. The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains. Upon convergence, $\hat{R}$, converges to 1. The number of model parameters for the four model fits are . The  maximum values of $\hat{R}$ across all model parameters for Models 1 throught 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004. Other diagnostics provided by the software (tree-depth, and divergent transitions) were clean. Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.

@@

A recently published \texttt{R} package, \texttt{loo}\cite{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent. It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\frac{\sum_{i=1}^S w_{i,s} p(t_{g,i})}{\sum_{s=1}^S w_{i,s}}$. In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result. For details, refer to \citet{vehtari}.

@@

Having chosen Model 4 as the best of the candidate models we considered, we would like to assess whether our model provides an adequate fit to the observed data. One way to do this is to draw replicate data sets from the posterior predictive distribution. This is a general approach recommended by \citet{gelman1996postpred} for models where no classical goodness-of-fit test is available. In the hierarchical setting, there can be different formulations of the predictive distribution. In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model. We choose to focus on the populations for which we have data collected.

@@

While TTF is clearly important, due to the
anticipation of advancements in technology, we can expect that the
relative value of computer hardware will depreciate rather quickly. In simple terms, given two hard drives, I would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all. To account
for depreciation, rather than using TTF as the metric for drive-model comparison, we use the value at replacement. The IRS considers
computer hardware to be ``5-year'' equipment\cite{f4562}. Using``the declining
balance method'' the rate of depreciation is 40\% per year.

@@

\appendix
\section{Truncation adjusted nonparametric estimate of lifetime}
\label{sec:trunc-adj}
We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
\cite{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
$t_{d,\text{min}}^L$, the shortest reported running time of all units
of drive-model $d$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by by Meeker and Escobar  in Chapter 11 of Reliability Analysis \cite{meeker}.  For each drive-model we select
$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
survived up to $t_{d,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by

@@

\pagebreak
\section{Posterior predictive graphical checks}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-1.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-2}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-3.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-4.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-5.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-6.pdf}
\end{figure}
\clearpage
% \bibliographystyle{plainnat}
\bibliographystyle{apalike}
% \bibliographystyle{plain}
\bibliography{sample}

>>>>

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep{nelson}.  A common assumption in series systems is the time to failure for each risk is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as ``masking''.\\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \citep{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et al. performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure \citep{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \citep{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \citep{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

@@

\section{Motivating Example}
\subsection{Backblaze hard drive data}
\label{sec:Data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is permanently removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its data publicly available through their website \citep{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.  No other analysis or modeling of the failure data is provided.  Backblaze does, however, encourage others to further analyze its data.

@@

\subsection{Single Drive-Model Analysis}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, drive-model 14.  The Kaplan-Meier estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures till about 10,000 hours, increases from about 10,000 to 20,000 hours, and then starts to gradually flatten out.  In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.  Therefore, there is both an empirical and theoretical justification to apply the GLFP model.

@@

Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \citep{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a detailed description).

In Figure 3 we plot the Kaplan-Meier adjusted cdf for drive model 14 on Weibull paper.  The solid stepwise curve corresponds to the observed failures.  Censored drives are not plotted.  Standard error bands are again calculated using Greenwood's method \citep{green}.  Also plotted is the median of the posterior draws for the GLFP model along with 90\% pointwise credible interval bands.  The non-parametric and parametric model align up quite closely.  Moreover, evidence of at least two failure modes is seen in the acceleration of the hazard rate between 8000 and 20,000 hours.  Therefore, a single Weibull model is clearly not flexible enough to characterize this lifetime distribution. The discrepancy in the right tail of the distribution appears to be real. We don't provide a solution to this problem here. Certainly, if there are clearly more that two well-separated failure modes, the GLFP will not fit the data well. However, we consider it satisfactory if we can provide a good approximation to the observed lifetime distribution over the bulk of the product life. In practical terms, the right tail is of less importance for electronic goods, due to obsolescence.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan} \citep{rstan} package in {\tt R} \citep{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\citep{betancourt}. HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density. This is done to reduce autocorrelation and efficiently explore the posterior. Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1,2 and 3 and 16 chains were run for Model 4. The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains. Upon convergence, $\hat{R}$, converges to 1. The number of model parameters for the four model fits are . The  maximum values of $\hat{R}$ across all model parameters for Models 1 throught 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004. Other diagnostics provided by the software (tree-depth, and divergent transitions) were clean. Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.

@@

A recently published \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent. It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $\log [p(t_{g,i}|t_{-(g,i)})]$ with $\frac{\sum_{i=1}^S w_{i,s} p(t_{g,i})}{\sum_{s=1}^S w_{i,s}}$. In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result. For details, refer to \cite{vehtari}.

@@

Having chosen Model 4 as the best of the candidate models we considered, we would like to assess whether our model provides an adequate fit to the observed data. One way to do this is to draw replicate data sets from the posterior predictive distribution. This is a general approach recommended by \cite{gelman1996postpred} for models where no classical goodness-of-fit test is available. In the hierarchical setting, there can be different formulations of the predictive distribution. In the present context, one could predict the lifetime of a new unit of a previously observed drive-model or a new drive-model. We choose to focus on the populations for which we have data collected.

@@

While TTF is clearly important, due to the
anticipation of advancements in technology, we can expect that the
relative value of computer hardware will depreciate rather quickly. In simple terms, given two hard drives, I would prefer to have both of them work for five years than for one to work for ten years and the other to not work at all. To account
for depreciation, rather than using TTF as the metric for drive-model comparison, we use the value at replacement. The IRS considers
computer hardware to be ``5-year'' equipment \citep{f4562}. Using``the declining
balance method'' the rate of depreciation is 40\% per year.

@@

\appendix
\section{Truncation adjusted nonparametric estimate of lifetime}
\label{sec:trunc-adj}
We first start with a non-parametric estimate of the empirical cdf for each drive-model using the Kaplan-Meier estimator
\citep{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator for drive-model $d$, denoted by
$\widehat{F_d(t)}_{KM}$, is conditional on survival up to
$t_{d,\text{min}}^L$, the shortest reported running time of all units
of drive-model $d$ for which records are available. To produce
unconditional estimates, we adapt the adjustment method outlined by by Meeker and Escobar  in Chapter 11 of Reliability Analysis \citep{meeker}.  For each drive-model we select
$t_{d,\text{min}}^L$, the smallest left truncated time in the sample.
By sampling from the full posterior distribution, since
$\Pr(T>t_{d,\text{min}}^L|\theta_d)$ (the probability a hard drive has
survived up to $t_{d,\text{min}}^L$) is a function of the model
parameters, we can easily compute its posterior median,
$\widehat{A}_{\text{med}} = \widehat{Pr}(T>t_{d,\text{min}}^L|\theta_d)$. We compute the adjusted estimate by

@@

\pagebreak
\section{Posterior predictive graphical checks}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-1.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-2}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-3.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-4.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-5.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-6.pdf}
\end{figure}
\clearpage
%\bibliographystyle{plainnat}
\bibliographystyle{asa}
% \bibliographystyle{plain}
\bibliography{sample}
